Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-10 write-ups/PLDI/paper.tex
Chat with Tyler and Hugues.

\begin{lstlisting}
__kernel connect_four(__global uchar *base_board,
                      const int lookahead,
                      __global atomic_int *best_move,
                      __global int *all_done,
                      __global Node *tree,
                      __global Task *pools,
                      __global Mutex *pool_m)
{
  // init
  if (get_global_id(0) == 0) {
    init_first_tasks(tree, pools);
  }
  global_barrier();

  // main loop
  while (true) {
    // obtain a task, or terminate if all work is done
    int pool_id = get_group_id(0);
    Task *t = pop_or_steal(pools, pool_m, pool_id);
    if (t == NULL) {
      if (*all_done) { return; } else { continue; }
    }

    // perform task, may create new tasks
    Node *node = tree[t->node_index];
    value = node_value(base_board, node);
    if (winning(value) || node->level == lookahead) {
      // we reached a leaf
      update_parent_value(node, tree, best_move, all_done);
    } else {
      // create_children() pushes or donates new tasks
      create_children(node, tree, pools, pool_m, pool_id);
    }
  }
}
\end{lstlisting}

@@

\lstset{basicstyle=\tt}
\begin{lstlisting}
  $\globalbarrier()$;
  if($\getgroupid{0}$ != 0) $\offerfork$();
  $\globalbarrier()$;
  $\offerkill()$;
  $\globalbarrier()$;
\end{lstlisting}
\lstset{basicstyle=\scriptsize\tt}

@@

\ADComment{Explain kernel merge tool.}

@@

@@

@@

To abstractly account for the interaction between global barrier
synchronization and the GPU memory model, we assume a function
$\mathsf{sync}$ that is applied to a global state.  This returns a new
global state in which the shared state and thread local states have
been updated to reflect the fact that all stores to memory and loads
from memory to thread-local storate have completed.

@@

\[
\inferrule{
\forall j \;.\; w_M(j) = (l_j, \offerkill();\mathit{ss})
\\
$M > 0$
}
{
(\sigma, (\dots, w_{M-1}, w_M, \bot, \dots, \bot)) \rightarrow (\sigma, (\dots, w_{M-1}, \bot, \bot, \dots, \bot))
}
(\textsc{Kill})
\]

@@

\[
\inferrule{
\forall i \;.\;\forall j\;.\;w_i(j) = (l_{i,j}, \globalbarrier();\mathit{ss})
\\
\forall i \;.\;\forall j\;.\;w_i'(j) = (l_{i,j}, \mathit{ss})
}
{
(\sigma, (w_1, \dots, w_M, \bot, \dots, \bot)) \rightarrow \mathsf{sync}(\sigma', (w_1', \dots, w_M', \bot, \dots, \bot))
}
(\textsc{Barrier})
\]

@@

\[
\inferrule{
\forall i \;.\;\forall j\;.\;w_i(j) = (l_{i,j}, \resizingglobalbarrier();\mathit{ss})
\\
\forall j\;.\;w_0'(j) = (l_{0,j}, \globalbarrier(); \offerfork(); \globalbarrier(); \globalbarrier();\mathit{ss})
\\
\forall i \neq 0 \;.\;\forall j\;.\;w_i'(j) = (l_{i,j}, \globalbarrier(); \globalbarrier(); \offerkill(); \globalbarrier();\mathit{ss})
}
{
(\sigma, (w_1, \dots, w_M, \bot, \dots, \bot)) \rightarrow (\sigma, (w_1', \dots, w_M', \bot, \dots, \bot))
}
(\textsc{Resizing-Barrier})
\]

>>>>

\ADComment{Points to make are: there is spinning, either on mutex implementation or main loop, so fairness is required.  Note OpenCL features: kernel keyword, global qualifier, ``get'' functions.  Note that pop\_or\_steal is executed by all, but internally the master manipulates the queue.  t is a single task, but is chunky enough that all threads can process part of it.}

\begin{lstlisting}
kernel work_stealing(global Task *pools) {
  while (more_work(pools)) {
    int pool_id = get_group_id();
    Task *t = pop_or_steal(pools, pool_id);
    if (t) {
      // task processing may create more work
      process_task(t, pools, pool_id);
    }
  }
}
\end{lstlisting}

@@

\lstset{basicstyle=\tt}
\begin{lstlisting}
  $\globalbarrier()$;
  if($\getgroupid()$ == 0) $\offerfork$();
  $\globalbarrier()$;
  $\offerkill()$;
  $\globalbarrier()$;
\end{lstlisting}
\lstset{basicstyle=\scriptsize\tt}

@@

\ADComment{Explain kernel merge tool briefly.  Explain library briefly.}

\ADComment{Give most attention to interesting performance and safety issues, e.g.\ reverse ticket lock, efficient barrier, the query functions...}

\ADComment{Host-device interaction through SVM.  Make clear that this is probably the most aggressive use of OpenCL 2.0 ever.}

@@

\subsection{Vendor Implementation Defects}

\ADComment{Briefly rant about our pain.  Compiler hang, compiler crashes, deadlocks, defunct processes.  Caution: don't make it sound so terrible that you can't trust our results.  Gently acknowledge that our implementation may be imperfect.}

@@

\ADComment{Tyler and Hugues, please write.}

@@

To abstractly accounts for the interaction between global barrier
synchronization and the GPU memory model, we assume a function
$\mathsf{sync}$ that is applied to a global state.  This returns a new
global state in which the shared state and thread local states have
been updated to reflect the fact that all stores to memory and loads
from memory to thread-local storate have completed.

@@

\[
\inferrule{
\forall j \;.\; w_M(j) = (l_j, \offerkill();\mathit{ss})
\\
M > 0
}
{
(\sigma, (\dots, w_{M-1}, w_M, \bot, \dots, \bot)) \rightarrow (\sigma, (\dots, w_{M-1}, \bot, \bot, \dots, \bot))
}
(\textsc{Kill})
\]

@@

\[
\inferrule{
\forall i \;.\;\forall j\;.\;w_i(j) = (l_{i,j}, \globalbarrier();\mathit{ss})
\\
\forall i \;.\;\forall j\;.\;w_i'(j) = (l_{i,j}, \mathit{ss})
}
{
(\sigma, (w_1, \dots, w_M, \bot, \dots, \bot)) \rightarrow \mathsf{sync}(\sigma , (w_1', \dots, w_M', \bot, \dots, \bot))
}
(\textsc{Barrier})
\]

@@

\[
\inferrule{
\forall i \;.\;\forall j\;.\;w_i(j) = (l_{i,j}, \resizingglobalbarrier();\mathit{ss})
\\
\forall j\;.\;w_1'(j) = (l_{1,j}, \globalbarrier(); \offerfork(); \globalbarrier(); \globalbarrier();\mathit{ss})
\\
\forall i \neq 1 \;.\;\forall j\;.\;w_i'(j) = (l_{i,j}, \globalbarrier(); \globalbarrier(); \offerkill(); \globalbarrier();\mathit{ss})
}
{
(\sigma, (w_1, \dots, w_M, \bot, \dots, \bot)) \rightarrow (\sigma, (w_1', \dots, w_M', \bot, \dots, \bot))
}
(\textsc{Resizing-Barrier})
\]
